# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
**Load the data (i.e. read.csv())**

```{r, echo=TRUE}
rm(list=ls())   #Clear all objects

#Load libraries
require(dplyr)
require(quantmod)  
require(lattice)  

#Set directory and file vectors
working.dir <- "C://Education//Reproducible Research//Homework//Peer Assessment 1"
input.dir <- ".\\input\\"
output.dir <- ".\\output\\"
script.dir <- ".\\scripts\\"
input.file.name <- "activity.csv"

#Set working directory
setwd(working.dir)

#Load the data
activity.data <- read.csv(paste(input.dir, input.file.name, sep="\\"), stringsAsFactors = FALSE)
```
**Process/transform the data (if necessary) into a format suitable for your analysis**
```{r, echo=TRUE}
steps.per.day <- activity.data %.%
  filter(steps >= 0) %.%  
  select(date, steps) %.%
  group_by(date) %.%
  summarize(sum.steps = round(sum(steps), 1)) %.%
  arrange(date)
```


## What is mean total number of steps taken per day?
**Make a histogram of the total number of steps taken each day**

```{r fig.width=7, fig.height=6, echo=TRUE}
hist(c(steps.per.day$sum.steps), main = "Histogram of the Total Number of Steps Takens Each Day", xlab = "Steps Per Day")
```
**Calculate and report the mean total number of steps taken per day**
```{r, echo=TRUE}
options(scipen = 1, digits = 7)
round(mean(steps.per.day$sum.steps), 2)
```
**Calculate and report the median total number of steps taken per day**
```{r, echo=TRUE}
options(scipen = 1, digits = 7)
median(steps.per.day$sum.steps)
```


## What is the average daily activity pattern?
**Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**
```{r, echo=TRUE}
average.daily.pattern <- activity.data %.%
  filter(steps >= 0) %.%  
  select(interval, steps) %.%
  group_by(interval) %.%
  summarize(mean.steps = round(mean(steps), 1)) %.%
  arrange(interval)

plot(average.daily.pattern, type="l", main = "Average Daily Activity Pattern", xlab = "5-minute interval", ylab = "Avg # of steps taken across all days", col = "blue")
```

## Imputing missing values
**Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)**
```{r, echo=TRUE}
nrow(activity.data[is.na(activity.data$steps), ])
```
**Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.**  

I will impute the missing values by using the mean for that 5-minute interval.

**Create a new dataset that is equal to the original dataset but with the missing data filled in.**  
```{r, echo=TRUE}
activity.data.non.na <- filter(activity.data, is.na(steps) == FALSE)
activity.data.na <- filter(activity.data, is.na(steps) == TRUE)
temp <- inner_join(average.daily.pattern, activity.data.na, by = c("interval"))
activity.data.new <- rbind(activity.data.non.na, select(temp, steps = mean.steps, date, interval))
```
**Make a histogram of the total number of steps taken each day v2**  
```{r, echo=TRUE}
steps.per.day.v2 <- activity.data.new %.%
  #filter(steps >= 0) %.%  
  select(date, steps) %.%
  group_by(date) %.%
  summarize(sum.steps = round(sum(steps), 1)) %.%
  arrange(date)
```
```{r fig.width=7, fig.height=6, echo=TRUE}
hist(c(steps.per.day.v2$sum.steps), main = "Histogram of the Total Number of Steps Takens Each Day v2", xlab = "Steps Per Day")
```
**Calculate and report the mean total number of steps taken per day v2**
```{r, echo=TRUE}
options(scipen = 1, digits = 7)
round(mean(steps.per.day.v2$sum.steps), 2)
```
**Calculate and report the median total number of steps taken per day v2**
```{r, echo=TRUE}
options(scipen = 1, digits = 7)
round(median(steps.per.day.v2$sum.steps),2)
```
**Do these values differ from the estimates from the first part of the assignment?**
The mean stays the same while the median increases slightly (i.e. 0.01%) from the first part of the assignment.

**What is the impact of imputing missing data on the estimates of the total daily number of steps?**  
```{r, echo=TRUE}
options(scipen = 1, digits = 1)
total.daily.steps.1 <- round(summarize(steps.per.day, steps = sum(sum.steps)), 2)
total.daily.steps.2 <- round(summarize(steps.per.day.v2, steps = sum(sum.steps)), 2)
delta <- Delt(total.daily.steps.1[,c("steps")], total.daily.steps.2[,c("steps")])*100
```
**The impact of imputing missing data on the estimates of the total daily number of steps is to increase the total daily steps from `r total.daily.steps.1` to `r total.daily.steps.2` (i.e. `r delta`%.)**

## Are there differences in activity patterns between weekdays and weekends?  
**Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.**  
```{r, echo=TRUE}
activity.data.new <- mutate(activity.data.new, weekday.text = weekdays(as.Date(activity.data.new$date)), weekday.text.baseline = weekdays(as.Date(activity.data.new$date)))
activity.data.new[activity.data.new$weekday.text == "Saturday" | activity.data.new$weekday.text == "Sunday", c("weekday.text")] <- "weekend"
activity.data.new[activity.data.new$weekday.text != "weekend" , c("weekday.text")] <- "weekday"
activity.data.new$weekday.text <- as.factor(activity.data.new$weekday.text)
```

**Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).**  
```{r fig.width=7, fig.height=6, echo=TRUE}
average.daily.pattern.new <- activity.data.new %.%
  #filter(steps >= 0) %.%  
  select(interval, steps, weekday.text) %.%
  group_by(interval, weekday.text) %.%
  summarize(mean.steps = round(mean(steps), 1)) %.%
  arrange(interval)

xyplot(average.daily.pattern.new$mean.steps ~ average.daily.pattern.new$interval | average.daily.pattern.new$weekday.text, data = average.daily.pattern.new, type = "l", layout = c(1, 2), xlab = "Interval", ylab = "Number of steps")
```
